From Likes to Change: Assessing the Impact of Citizen Engagement on the European

### 2023-06-13 Article.Rmd
2

---
title: 'From Likes to Change: Assessing the Impact of Citizen Engagement on the European
  Commission''s Social Media Platforms'
author:
- name: Tanase Tasențe
  affiliation: '1'
  corresponding: yes
  address: Postal address
  email: "tanase.tasente@365.univ-ovidius.ro"
  role:
  - Equal contribution
  - Conceptualization
  - Data curation
  - Investigation
  - Formal analysis
  - Resources
  - Software
  - Visualization
  - "Writing - Original Draft Preparation"
  - "Writing - Review & Editing"
- name: "Cristian Opariuc-Dan"
  affiliation: '1,2'
  role:
  - Equal contribution
  - "Writing - Original Draft Preparation"
  - "Writing - Review & Editing"
  - Data curation
  - Formal analysis
  - Resources
  - Software
  - Visualization
- name: Cristina Dana Popescu
  affiliation: '1'
  role:
  - "Writing - Original Draft Preparation"
  - "Writing - Review & Editing"
  - Professional proofreading
shorttitle: Impact of Citizen Engagement on the European Commission's Social Media
  Platforms
output:
  html_document:
    df_print: paged
  word_document: papaja::apa6_docx
  pdf_document: papaja::apa6_pdf
authornote: |
  Department of Administrative and Social Sciences
abstract: "This study investigates the role of emotional resonance in social media
  communications from the European Commission (EC) and its impact on public participation
  across various platforms including Facebook, Instagram, Twitter, and YouTube. Findings
  indicate that emotionally resonant communications significantly enhance public engagement,
  with variations depending on the specific platform. Both positive and negative emotional
  resonance significantly impact participation, with the latter having a more pronounced
  effect. Our research highlights the critical importance of platform-specific strategies
  and the incorporation of emotional resonance in public communications to foster
  participation in policy dialogues. An analysis of the engagement rates and sentiment
  across platforms reveals unique tonal variations in the EC's discourse. The effect
  of emotional resonance on public engagement is also found to be moderated by the
  choice of social media platform, with certain platforms more significantly moderating
  the relationship between emotional resonance and engagement than others. This study
  lays the groundwork for refining communication strategies within a multi-platform
  digital governance environment.\n"
keywords: "social media, public institutions, emotional resonance, public engagement,
  platform-specific strategies, digital governance"
wordcount: 4067 words in text body 1034 words in reference section
bibliography:
- "r-references.bib"
- references.bib
floatsintext: no
linenumbers: yes
draft: no
mask: no
figurelist: no
tablelist: no
footnotelist: no
documentclass: apa7
classoption: man
csl: apa.csl
affiliation:
- id: '1'
  institution: Ovidius University of Constanța, Romania
- id: '2'
  institution: University of Bucharest, Romania
header-includes: \usepackage{ragged2e}
---

@@

# Discussion
The present investigation embarked on a quest to explore public engagement across various social media platforms utilized by the European Commission, with an emphasis on the impact of emotional resonance within content. The role of social media in public institutions' communication strategies is pivotal [@Smith2015], necessitating a deep understanding of engagement dynamics.

The data supports our first hypothesis (H1), suggesting that communications rich in emotional resonance, whether of positive or negative valence, are correlated with enhanced public engagement. However, the ability of emotional resonance to elucidate engagement levels is negligible (0.16%). This corroborates the complex nature of engagement behaviours expounded in scholarly literature [@Dolan2016; @Dragseth2020; @Mirbagheri2019] and points towards the necessity for further examination of supplementary factors influencing public engagement.

Our second hypothesis (H2) posited the influence of the social media platform choice on the effect of emotional resonance on public engagement. The empirical evidence supports this, suggesting that different platforms may cater to varied user expectations and norms. These findings align with prior research that demonstrated platform-specific fluctuations in engagement patterns [@Voorveld2018], underscoring the need for the European Commission to acknowledge heterogeneous engagement patterns and tailor their communications accordingly.

The third hypothesis (H3) asserted that the social media platform would moderate the impact of emotional resonance on public engagement. Our data confirmed this assumption, revealing a statistically significant moderating effect. This suggests that platforms may shape the way the public interacts with emotionally resonant messages, emphasizing the importance of customizing such communications to align with each platform's norms.

This research has limitations due to the use of secondary data obtained from platforms, suggesting future studies could benefit from primary data collection. Moreover, this study did not associate Engagement Rate (ER) with significant concurrent events, potentially overlooking the impact of emotional reactions based on these events. Future studies could explore these associations to understand emotional impacts better, thus extending the existing knowledge in the field.

Additionally, the scope of our research was limited to the European Commission's use of four major social media platforms. Further studies could explore other platforms or public institutions to see if the observed patterns hold. Also, while we focused on emotional resonance, other aspects of communication such as tone, complexity, and framing could be investigated in future research.

In conclusion, this study underscores the significance of emotional resonance and platform-specific norms in driving public engagement. It suggests that forming an emotional connection with the audience through posts could amplify engagement levels. By understanding these dynamics, public institutions like the European Commission can refine their social media strategies, enhance citizen engagement, and bolster the efficacy of their public communication. Despite the limitations, the study provides a valuable starting point for future research into the multifaceted world of social media engagement.

>>>>

---
title: 'From Likes to Change: Assessing the Impact of Citizen Engagement on the European
  Commission''s Social Media Platforms'
author:
- name: Tanase Tasențe
  affiliation: '1'
  corresponding: yes
  address: Postal address
  email: "tanase.tasente@365.univ-ovidius.ro"
  role:
  - Equal contribution
  - Conceptualization
  - Data curation
  - Investigation
  - Formal analysis
  - Resources
  - Software
  - Visualization
  - "Writing - Original Draft Preparation"
  - "Writing - Review & Editing"
- name: "Cristian Opariuc-Dan"
  affiliation: '1,2'
  role:
  - Equal contribution
  - "Writing - Original Draft Preparation"
  - "Writing - Review & Editing"
  - Data curation
  - Formal analysis
  - Resources
  - Software
  - Visualization
- name: Cristina Dana Popescu
  affiliation: '1'
  role:
  - "Writing - Original Draft Preparation"
  - "Writing - Review & Editing"
  - Professional proofreading
shorttitle: Impact of Citizen Engagement on the European Commission's Social Media
  Platforms
output:
  pdf_document: papaja::apa6_pdf
  word_document: papaja::apa6_docx
  html_document:
    df_print: paged
authornote: |
  Department of Administrative and Social Sciences
abstract: "This study investigates the role of emotional resonance in social media
  communications from the European Commission (EC) and its impact on public participation
  across various platforms including Facebook, Instagram, Twitter, and YouTube. Findings
  indicate that emotionally resonant communications significantly enhance public engagement,
  with variations depending on the specific platform. Both positive and negative emotional
  resonance significantly impact participation, with the latter having a more pronounced
  effect. Our research highlights the critical importance of platform-specific strategies
  and the incorporation of emotional resonance in public communications to foster
  participation in policy dialogues. An analysis of the engagement rates and sentiment
  across platforms reveals unique tonal variations in the EC's discourse. The effect
  of emotional resonance on public engagement is also found to be moderated by the
  choice of social media platform, with certain platforms more significantly moderating
  the relationship between emotional resonance and engagement than others. This study
  lays the groundwork for refining communication strategies within a multi-platform
  digital governance environment.\n"
keywords: "social media, public institutions, emotional resonance, public engagement,
  platform-specific strategies, digital governance"
wordcount: 4067 words in text body 1034 words in reference section
bibliography:
- "r-references.bib"
- references.bib
floatsintext: no
linenumbers: yes
draft: no
mask: no
figurelist: no
tablelist: no
footnotelist: no
documentclass: apa7
classoption: man
csl: apa.csl
affiliation:
- id: '1'
  institution: Ovidius University of Constanța, Romania
- id: '2'
  institution: University of Bucharest, Romania
header-includes: \usepackage{ragged2e}
---

@@

# Discussion
The current study embarked on an investigative journey to examine public engagement across a range of social media platforms utilized by the European Commission, laying specific emphasis on the impact of emotional resonance within disseminated content. The vital role of social media in communication strategies of public institutions has been widely recognized [@Smith2015], accentuating the necessity for an in-depth comprehension of the dynamics of engagement.

Our collected data validates our primary hypothesis (H1), implying that communications imbued with emotional resonance, regardless of its positive or negative polarity, are associated with augmented public engagement. Nevertheless, the capacity of emotional resonance to illuminate engagement levels remains marginal (0.16%). This finding is consistent with the complex nature of engagement behaviours outlined in existing academic discourse [@Dolan2016; @Dragseth2020; @Mirbagheri2019], thereby underscoring the need for further scrutiny of additional factors influencing public engagement.

Our secondary hypothesis (H2) proposed that the choice of the social media platform would significantly influence the effect of emotional resonance on public engagement. The empirical evidence corroborates this assertion, indicating that different platforms potentially cater to diverse user expectations and conventions. This alignment with prior research revealing platform-specific variations in engagement patterns [@Voorveld2018] underscores the necessity for the European Commission to recognize and adapt to these heterogeneous engagement patterns when designing their communication strategies.

The tertiary hypothesis (H3) posited that the chosen social media platform would moderate the impact of emotional resonance on public engagement. The acquired data affirmed this premise, demonstrating a statistically significant moderating effect. This indicates that the characteristics of individual platforms may shape the public’s interaction with emotionally resonant messages, highlighting the need to tailor such communications to align with the norms specific to each platform.

This research is constrained by its reliance on secondary data obtained from platforms, suggesting that future studies could potentially reap benefits from primary data collection methodologies. Furthermore, this study did not associate Engagement Rate (ER) with significant simultaneous events, potentially overlooking the effect of emotional reactions provoked by these events. Future investigations could probe these associations, thereby enhancing our understanding of emotional impacts and expanding the existing body of knowledge in this domain.

Additionally, the scope of this research was confined to the European Commission's use of four primary social media platforms. Future inquiries could potentially explore other platforms or public institutions to determine the universality of the observed patterns. Furthermore, while our focus lay primarily on emotional resonance, future research could delve into other aspects of communication, such as tone, complexity, and framing.

To conclude, this study underscores the imperative role of both emotional resonance and platform-specific norms in propelling public engagement on social media. The ability to establish an emotional connection with the audience, made possible through well-conceived and executed posts, emerges as a powerful tool for amplifying engagement levels. Such engagement is not merely about expanding the reach or visibility of public institutions like the European Commission, but is a mechanism for fostering a dynamic and participatory relationship between these institutions and the citizens they serve.

Understanding these dynamics enables public institutions to tailor their communication strategies more effectively. By acknowledging the varying norms and expectations tied to different social media platforms, institutions can ensure that their messages resonate more strongly with their intended audiences, ultimately enhancing the impact and penetration of these messages. The choice of platform is not incidental but plays a significant role in how emotional resonance influences engagement, urging public institutions to be thoughtful and strategic in their selection and use of these platforms.

Moreover, by refining their social media strategies based on these findings, public institutions have the opportunity to foster greater citizen engagement. Higher engagement can lead to increased public awareness, enhanced understanding, and potentially even behavioural change - critical goals for public communication strategies. It can also help public institutions gain insight into public sentiment, thus providing valuable feedback that can be used to improve policies and services.

In addition, the insights derived from this study can help public institutions increase the effectiveness of their public communication. By leveraging emotional resonance and considering platform-specific norms, these institutions can design messages that are more likely to engage citizens and generate a meaningful response. This has the potential to transform the landscape of public communication, making it a more interactive, responsive, and impactful domain.

Despite its limitations, this study provides an invaluable point of departure for further research into the intricate world of social media engagement. It opens up avenues for exploring other influential factors and for deepening our understanding of the relationship between emotional resonance, platform norms, and public engagement. These insights could be crucial in helping public institutions navigate the complex dynamics of social media engagement, enhancing their ability to communicate effectively with the public, and ultimately, improving their service to society.
